---
title: "Classcification Report for Eating and Health Module"
author: "Group Heheda"
date: "12/5/2020"
output: html_document
---

### Introduction

  Nowadays, as the socioeconomoic level of people' lives develops, more and more attention are paid at enhancing personal health and lifestyles. Eating health, as an indispensable component of individual's biological maintainence, plays a vital role in strengthening personal immunity and preventing chronic diseases, including regular eating habits and nutritional diets. Addressed in "Eating and Health Module" (2019, Economic Research Service), eating patterns are strongly affected by people's economic status and behaviors, including household income, engagements in physical activities, food preparation styles, accessibilities to grocery stores, and etc. From the aspect of public health and policy, understanding how these indicators affect personal eating health can assist the government to develop effective and efficient food and nutrition assistance programs, increasing stable and harmonized relationship between people and society. This project, starting from these beneficial standspoint, explores impacts of  different indicators on personal eating health.
  
### Data

#### Dataset Information
  The dataset is obtained and downloaded from Economic Research Service of United States Department of Agriculture, named "Eating and Health Module". The dataset is case-specific repsondent file, including general personal biolgical information (height, weight), household information (income, engaged government assistance programs), eating habits (fast food consumption, drink habits, food preparation), and physcial activities of repsondents. The whole data collection was conducted by Questionnaire and Survey, based on answers of voluntary respondents. We mainly started with exploring sample selection of repsondents via several descriptive statistics on their general health information. Then we analyzed associations or distributions of internal or external indicators for understanding their impacts on personal eating health by using persoanl BMI value as the parameter, including accessibilities of grocery stores, conditions of household income, frequency of fast food consumptions and engagements of physical activities. Analysis would include different association plots and multi-linear regression model. We aso established the hypothesis testing for measuring similarity or difference resided between primary and secondary eating behaviors of respondents.

```{r, include=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(pastecs)
library(plotly)
library(goftest)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
raw_eathealth = read_csv("ehresp_2014.csv")
```

#### Tidy Process
  The raw dataset contains `r nrow(raw_eathealth)` observations and `r ncol(raw_eathealth)` variables. We selected 10 variables for our project. Variables consists of categorical variables, logical variables and continuous with the validation rule of reporting positive or specified positive value. However, because the dataset is totally based on willingness of voluntary respondent via Questionnaire, there are a large amount of invalid data entries, like negative values as the repsondent was not willing to answer certain questions. Also, for multiple continous variables, like frequency of fast food consumptions or frequency of exercises, data mainly focused on certain integers, resulting that continuous variables are more likely categorical variables. 
  
  Because the original dataset contains lots of categorical variables, we firstly recode all categorical variable into corresponding categories following the codebook provided along with the dataset. And for excluding invalid values, the original raw dataset was tidied into desired one based on different statistical analysis. 

### Statistical Analysis

#### Descriptive Statistics for Respondents

# Income level.
```{r}
income_tbl = 
  raw_eathealth %>%
  count(tucaseid) %>%
  mutate(
    erincome = factor(raw_eathealth$erincome, levels = c(-1, 1, 2, 3, 4, 5), labels = c("Blank", "Income > 185% of poverty threshold", "Income <= 185% of poverty threshold", "130% of poverty threshold < Income < 185% of poverty threshold", "Income > 130% of poverty threshold", "Income <= 130% of poverty threshold"))
  ) %>%
  plot_ly(x = ~erincome, y = ~n, color = ~erincome, type = "bar")

income_tbl
```

The majority of the study objects are on the highest income level - their income is greater than 185% of poverty threshold. The income level with the second largest study population come from people who are below or equal to 130% of poverty threshold. Only 36 study subjects have income greater than 130% of poverty threshold.

# BMI, height, and weight.
```{r}
raw_eathealth %>%
   select(erbmi, euhgt, euwgt) %>%
   filter(erbmi > 0, euhgt > 0, euwgt > 0) %>%
   summary() %>%
   format(scientific = F, digits = 2)
```
Among all study objects, the mean height is 66.69 inches, and the mean weight is 176.3 pounds.

The mean BMI is 27.77, and third quantile is 30.7, while the maximum BMI is 73.6. We suspect there might be some BMI outliers on the upper end.

```{r}
raw_eathealth %>%
  filter(erbmi > 0, euhgt > 0, euwgt > 0) %>%
  ggplot(aes(x = "", y = erbmi)) + 
  geom_boxplot()
```

The boxplot did show that there are more outliers on the upper end than on the lower end of BMI.

#### Association Analysis for different Indicators vs. Eating Health
Here in this section, want to examine the asscociation between external factors and BMI.Sometimes we are intersted in the category that the BMI is refer to. .According to the WHO BMI standard, we categorize BMI into four categories: "Underweight","Normal","Overweight","Obese", then we want to comare the proportion of each BMI status among indicator categories.Therefore, according to distribution of the data, for indicatorsincome and household conditions, and we use categorized BMI

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height = 12, fig.width = 16)
install.packages("plotrix")
library(ggplot2)
library(flexdashboard)
library(tidyverse)
library(dplyr)
library(plotly)
library(plotrix)

theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r include=FALSE}
raw_eathealth_2 = read.csv("ehresp_2014.csv") %>%
  mutate(
    tucaseid = as.factor(tucaseid)
  ) %>%
  select(tucaseid, starts_with(c("er", "eu")), -erhhch)

```


```{r include=FALSE}
external_indicators_two = raw_eathealth_2 %>%
  select(erbmi, eugenhth, erincome, eustores,erspemch) %>%
  filter_all(all_vars(. > 0)) %>%
  rename("physical_health" = "eugenhth") %>%
  mutate(
    physical_health = factor(physical_health, labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")),
     erincome = factor(erincome, labels = c("Income > 185% of poverty threhold", "Income <= 185% of poverty threshold", "130% < Income < 185% of poverty threshold", "Income > 130% of poverty threshold", "Income <= 130% of poverty threshold")),
    eustores = factor(eustores, labels = c("Grocery Store", "Supercenter", "Warehouse Club", 
                                           "Drugstore or Convenience Store", "Some Other Place")),
    erspemch = factor(erspemch, labels = c("Partner's work time decrease", "Partner's work time increase", "Partner lose job","Partner find job", "Not change" ))
       )%>%
rename("bmi"="erbmi", "income"="erincome","place_get_groceries" ="eustores","change_in_work_of_sponsor" = "erspemch")
```
#categorize BMI
```{r}
external_category_crude =
raw_eathealth_2 %>%
  select(erbmi, eugenhth, erincome, eustores,erspemch) %>%
  filter_all(all_vars(. > 0)) %>%
  rename("physical_health" = "eugenhth") %>%
  mutate(
    physical_health = factor(physical_health, labels = c("Excellent", "Very Good", "Good", "Fair", "Poor")),
     erincome = factor(erincome, labels = c(">185% ", "<=185% ", "130%-185% ", ">130% ", "<=130%")),
    eustores = factor(eustores, labels = c("Grocery Store", "Supercenter", "Warehouse Club", 
                                           "Drugstore or Convenience Store", "Some Other Place")),
    erspemch = factor(erspemch, labels = c("Partner's work time decrease", "Partner's work time increase", "Partner lose job","Partner find job", "Not change" ))
       )%>%
rename("bmi"="erbmi", "income_threhold"="erincome","place_get_groceries" ="eustores","change_in_work_of_spouse" = "erspemch") %>%
mutate(bmi_new =cut(bmi, breaks=c(-Inf,18.5, 25, 30, Inf), labels=c("Underweight","Normal","Overweight","Obese")))
```

# Plot for Distribution by Income and BMI
```{r}
counts <- table(external_category_crude$bmi_new, external_category_crude$income)
barplot(counts, main="Distribution by Income and BMI",
  xlab="Income level", col=c("#8FB78F","#2A6E93","#6495ED","#ff8754"),
  legend = rownames(counts),beside=TRUE)
```

From the plot, we could see that the higher-income population has larger proportion of normal weighted population and lower proprotion of Obese population than in the lower-income groups. One thing worth to mention is  that the proportion of obese populaion is even higher than that of normal-weighted person among the lowest income group, which give us public health implications that we need to consider the low-income population as vulerable population for obesity.


# Plot for Distribution by spouse conditions and BMI
```{r}
external_categ_house = external_category_crude %>%
group_by(bmi_new,change_in_work_of_spouse)%>%
mutate(count_house=n())%>%
count(bmi_new)
  
```

```{r}
house=
ggplot(external_categ_house) + geom_col(aes(x = 1, y = n, fill =bmi_new), position = "fill") +coord_polar(theta = "y") + facet_wrap(~ change_in_work_of_spouse) 
house
```
We are interested in this area is because we concern whether spouse's working status would actually influence the health status. Interestingly, the proportion of normal population among partner's working time increase are approximaterly the same as that among partner's lose job. Also, the proportion of normal weightes population among partner's lose job a are significantly higher than that among partner's find  job.In addition, we could not see a trend of increase or decrease of the proportion of neither normal weighted nor obese population among these five household conditions. Therefore, it seems like spouse's working status will not significantly influence BMI.

# Plot for preferenece in grocery store 
```{r}
plot_stor_bmi_one=
external_indicators_two %>%
ggplot(aes(x=bmi,fill = place_get_groceries))+geom_density(alpha=0.5)+facet_grid(place_get_groceries~.)+scale_x_continuous( breaks = seq(10,60,3),limits =c(10,60))+labs( title = "Distribution of BMI for different places get groceries", x = "BMI", y="Density")+theme(legend.position = "bottom", plot.title = element_text(vjust = 5),panel.spacing = unit(1, "lines"))+scale_fill_manual(values=c("#6495ED","#8FB78F","#56B4E9","66cc99","#FF8754"))
plot_stor_bmi_one
```
From the density plot, we could se that the distribution for the BMI are approximatelu the same among people who shops in Crocerystore, Supercenter and warehouse club. However, for people who Consume products in Drugstore or convenience store, the BMI are relative higher that shoppers with other shopping-preferences, one thing need to mention that people who shoppers in other places looks relatively healthier, we could inverstigate where they shop or why they made this decision, and explore this in the future. 



#### Multi-regression Model for BMI vs. Frequency of Fast Food Consumption and Exercises

  As the rhythm of modern lifestyles expedites, fast food gradually becomes the mainstream of daily meals, mainly consists of high carbohydrates, salts, fats and low celluloses. Long term intake of fast food has been correlated to multiple chronic diseases, such as diabetes, poor nutrition and obesity. Unlike fast food consumption, exercise plays a key role in maintaining and strengthening good physical health. Generally, both two indicators would show different kinds of associations with personal BMI in magnitudes and signs.We built a multi-regression model for exploring association between BMI and frequency of fast food consumption with incorporation of frequency of exercises.

```{r, echo = FALSE, fig.height=8, fig.width=10, warning=FALSE, message=FALSE}

library(plotly)
raw_eathealth = read_csv("ehresp_2014.csv") %>%
  mutate(
    tucaseid = as.factor(tucaseid)
  ) %>%
  select(tucaseid, starts_with(c("er", "eu")), -erhhch)

internal_indicators = raw_eathealth %>%
  select(erbmi, eufastfdfrq, euexfreq) %>%
  filter_all(all_vars(. > 0))

fig_data = internal_indicators %>%
  pivot_longer(
    eufastfdfrq : euexfreq,
    names_to = "indicator",
    values_to = "value"
  ) %>%
  mutate(
    indicator = recode(indicator, "eufastfdfrq" = "Fast Food Consumption",
                       "euexfreq" = "Exercise"),
    indicator = factor(indicator, levels = c("Fast Food Consumption", "Exercise")))

mlr_plot = fig_data %>%
  lm(erbmi ~ value * indicator, data = .) 
  
fig_data %>%
  mutate(
    fit_value = fitted(mlr_plot)
  ) %>%
  plot_ly(x = ~ value, y = ~erbmi, frame = ~ indicator, color = ~ indicator, type = "scatter") %>%
  add_lines(x = ~ value, y = ~ fit_value, frame = ~ indicator, mode = "lines") %>%
  layout(title = "Scatter plots of BMI vs. Frequency of Fast Food Consumption and Exercise",
         xaxis = list(title = "Frequency"), yaxis = list(title = "BMI")) %>%
  animation_opts(2500, easing = "linear", redraw = T) %>%
  animation_slider(hide = F)

summ_mlr_plot = summary(mlr_plot)

```
  
  From the scatter plot of BMI vs. Frequency of Fast Food Consumption, originally continuous variables are shown approximately to categorical variables, which might be caused by the way of data collection: Questionnaire. Generally, the regression line of BMI vs. frequency of fast food consumption is positive. As the consumption of fast food increases, the value of BMI would also increase, indicating that fast food consumption relates to decreased eating health. The regression line of BMI vs. frequency of exercise is negative, which means higher frequency of exercise would result in lower BMI. However, according to the R-squared of `r summ_mlr_plot[["r.squared"]]`, the model need further adjustment for detecting associations between BMI and two kinds of frequencies. So, multi-linear regression model incorporating both indicators was built.
```{r, echo=FALSE, fig.height=8, fig.width=10, warning=FALSE, message=FALSE}

mlr_model = lm(erbmi ~ eufastfdfrq + euexfreq, data = internal_indicators)
summ_mlr = summary(mlr_model)

par(mfrow = c(2, 2))
plot(mlr_model)

```  
  
  Multiregression model of BMI vs. Frequency of Fast Food Consumption and Exercise showed that the estimated coefficient for fast food consumption is `r round(mlr_model[["coefficients"]][["eufastfdfrq"]], 4)` with p.value of `r round(summ_mlr[["coefficients"]][11], 4)`and for exercise is `r round(mlr_model[["coefficients"]][["euexfreq"]], 4)` with p.value of `r summ_mlr[["coefficients"]][12]`. R-squared: `r round(summ_mlr[["r.squared"]], 4)` 
  Both p.value prove the statistical significance of two coefficients in the multiregression model. Low R-squared value allows more residual and outlier tests for the model. Obviously, residuals for the model are deviated from normal distribution. Both scatter plots and Leverage plots show that there are multiple outliers resided in the model, affecting the fitted model for the regression of BMI vs. frequency of food consumption and exercise.
  Adjustments are performed by controlling existed outliers.

```{r, echo=FALSE,  fig.height=8, fig.width=10, warning=FALSE, message=FALSE}
adjust_model = internal_indicators %>%
  mutate(
    rstand = rstandard(mlr_model)
  ) 

outliers = adjust_model %>%
  filter(abs(rstand) > 2.5)

mlr_adjust = adjust_model %>%
  filter(abs(rstand) < 2.5) %>%
  lm(erbmi ~ eufastfdfrq + euexfreq, data = .)
mlr_adjust_summ = summary(mlr_adjust)

qqnorm(resid(mlr_adjust), xlab = "Expected Value", ylab = "Residual", 
       main = "QQ PLot of Residual for Adjusted-MLR")
qqline(resid(mlr_adjust))
```

  `r nrow(outliers)` outliers are excluded by "internally studentized residuals".  Comparing with unadjusted model, adjusted multiregression model with exclusion of outliers showed lower estimated slope of `r round(mlr_adjust[["coefficients"]][["eufastfdfrq"]], 4)`for linear regression BMI vs. Frequency of food consumption with higher statistical significance proved from lower p.value of `r round(mlr_adjust_summ[["coefficients"]][11], 4)`. QQ plot presented that exclusion of outliers made the residuals more approaching to normal distribution. Regression of BMI vs. exercise presented higher estimated slope of `r mlr_adjust[["coefficients"]][["euexfreq"]]`. 
  R-squared: `r round(mlr_adjust_summ[["r.squared"]], 4)` with increment of `r round(mlr_adjust_summ[["r.squared"]], 4) - round(summ_mlr[["r.squared"]], 4)`
  Adjusted R-squared: `r round(mlr_adjust_summ[["adj.r.squared"]], 4)` with increment of `r round(mlr_adjust_summ[["adj.r.squared"]], 4) - round(summ_mlr[["adj.r.squared"]], 4)`
  
  In general, due to Questionnaire way used on data collection, there would be bias existed while respondents answered frequency of fast food consumption and exercises. From the multiregression model, we can conclude that exercises affects BMI in the positive way which improving personal phyiscal health as the frequency of engaging activities increases. Even though the estimated coefficient of BMI vs.Frequency of fast food consumption is only `r round(mlr_adjust[["coefficients"]][["eufastfdfrq"]], 4)`, it is consistent with what we expected before establishing the model that the physical health will decreases as the consumption of fast food increases. In other words, eating health is negatively impacted by fast food consumption but positively impacted by exercises.
  
#### Hypothesis Testing

We want to test whether respondents who spent longer time in primary eating or drinking would also be those who spent longer time in secondary eating.

H0: mean primary eating time = mean secondary eating time.
H1: mean primary eating time =\ mean secondary eating time.

Statistical method: paired t-test. 

```{r}

prim_sec_eat = raw_eathealth %>%
  select(erbmi, ertpreat, ertseat) %>%
  filter_all(all_vars(. > 0))

prim_sec_eat =
  prim_sec_eat%>%
  mutate(
    diff_time = ertpreat - ertseat,
    diff_time_num = as.numeric(diff_time)
  )


ad.test(prim_sec_eat$diff_time_num)
``` 
According to Anderson-Darling normality test, since p-value is smaller than 0.05, we reject null hypothesis and conclude that the difference between primary eating time and secondary eating time is normally distributed. Therefore, we can use paired t-test.

```{r}
t.test(prim_sec_eat$ertpreat, prim_sec_eat$ertseat)
```
According to paired t-test, since p-value is smaller than 0.05, we reject null hypothesis and conclude that respondents do not spend same amount of time on primary eating and secondary eating. We are 95% confident that respondents spend 31 to 35 minutes longer on primary eating than on secondary eating. 

### Conclusion
