---
title: "Hypothesis Testing"
author: "Yue Chen"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

library(tidyverse)
library(dplyr)
library(plotly)
library(goftest)

```

# Tidy Dataset
```{r}

raw_eathealth = read.csv("ehresp_2014.csv") %>%
  mutate(
    tucaseid = as.factor(tucaseid)
  ) %>%
  select(tucaseid, starts_with(c("er", "eu")), -erhhch)

```

## C. Hypothesis Testing

We want to test whether respondents who spent longer time in primary eating or drinking would also be those who spent longer time in secondary eating.

H0: mean primary eating time = mean secondary eating time.
H1: mean primary eating time =\ mean secondary eating time.

Statistical method: paired t-test. 

```{r}

prim_sec_eat = raw_eathealth %>%
  select(erbmi, ertpreat, ertseat) %>%
  filter_all(all_vars(. > 0))

prim_sec_eat =
  prim_sec_eat%>%
  mutate(
    diff_time = ertpreat - ertseat,
    diff_time_num = as.numeric(diff_time)
  )


ad.test(prim_sec_eat$diff_time_num)
``` 
According to Anderson-Darling normality test, since p-value is smaller than 0.05, we reject null hypothesis and conclude that the difference between primary eating time and secondary eating time is normally distributed. Therefore, we can use paired t-test.

```{r}
t.test(prim_sec_eat$ertpreat, prim_sec_eat$ertseat)
```
According to paired t-test, since p-value is smaller than 0.05, we reject null hypothesis and conclude that respondents do not spend same amount of time on primary eating and secondary eating. We are 95% confident that respondents spend 31 to 35 minutes longer on primary eating than on secondary eating. 